Automated topic naming to support analysis of software maintenance activities

### 2010-08-21 icse/icse2011.tex
generating the data
minor edits

% why are topics useful
Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifacts themselves.
Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}. 
Without the structured information which dashboards typically rely on, topics become a valuable and automatically extractable abstraction of software artifacts.
Topics abstract software artifacts by their subject matter, rather than their concrete representation. 
Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
The topics of development are relevant because they identify the distinct issues facing development at a particular time.

The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
% get neil's data mining point in there
To date, the results of topic analysis and concept location are project specific, and not generalizable.
% get the interpretation aspect
Much of the current research stops short of automatically interpreting the purpose of the topic. 
That is, extracted topics are usually manually named and labelled by those who read the topic words.

% topics are numerous?
Deriving such topics is difficult. 
Topic analysis produces % raw topics that are represented as
 unlabelled context-free word-lists or word distributions.
Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
This manual approach is impractical when one has to handle hundreds of different topics. To scale to meaningfully sized projects, we need automatic assistance to determine the topic label.
We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.

@@

% harp on data mining more
In general, the mining of software artifacts tends to be very project specific, yet NFRs are not. 
% I was told if you're going to use a paren, you don't need to say it.
There is a series of standards on NFRs, such as \cite{iso9126}, specifically intended to apply to projects of varying types.
These standards indicate our approach is possible.%, if perhaps not universally accepted).
% NFRs are prevalent in almost all software systems.
% In this paper we will show that NFR related topics occur quite frequently.
%XXXXXX this sentence is confusing
This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. 
In particular, we use software quality models to generalize topics across projects. 
This allows us to mine software systems with the intention of comparing their NFR-related development topics.
% in order to achieve generalizable, cross project data mining results.

@@

To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
We selected systems from the same application domain, so the functional requirements would be broadly similar. 
We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. 
The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
MySQL contains 320 thousand lines of C and C++ source code.  
Choosing an older version of these projects allows us to focus on projects which have moved into the maintenance phase of the software life-cycle.

@@

For the commit message data-sets of each project, we created an XML file which separated commits into 30 day periods. 
This size of period, 30 days, is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. 
For each period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis algorithm~\cite{Blei2003}, and recorded the topics of each period.
%For each project and for each period of messages a project, we input these messages 
%we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.

@@

We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
We set the number of topics to generate to $20$, because past experimentation, e.g. \cite{Hindle09ICSM}, showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics. 
Furthermore, more than $20$ topics quickly became infeasible for inspection.

@@

\textbf{Do label frequencies change over time?} -- In both projects the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
As discussed in the previous section, certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time horizon being shorter than an equivalent span for the MySQL product. %XXXXXXXXXXXXXXXXX what

@@

% para 4 summarize and hammer contribution
% NE not a fan of this phrasing -> % In conclusion, 
Previous topic analysis research produced topics that needed to be manually labelled, but topics without labels are difficult to interpret.
To improve this, we leveraged software engineering standards to produce a method of domain specific topic labelling.
Since the word-list technique is not project specific, it can be used to compare multiple projects.
We demonstrated multiple supervised and unsupervised methods of topic labelling, including multi-label learners.
We provided a case study and visualization of NFR related topics in MaxDB 7.500 and MySQL 3.23. 
During the case study we manually labelled hundreds of topics by hand and then used these annotations to validate the positive performance of topic labelling techniques.
The implication of our paper is that topics with labels are more useful than unlabelled topics. To furnish stakeholders with labelled topics, our technique provides many methods of automatic topic labelling.

@@

% Developers do more than merely observe these abstractions, however. They must also use them to assess progress and work with colleagues. To have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our development artifacts. Such labels can be used to compare progress or understand how well a project is meeting requirements. These labels then become the topics for development conversations (such as the stream of commentary on a bug report). One way to identify topics is to use the artifacts themselves to derive them. 
% \todo[inline]{it should be labels for topics not artifacts, I think the artifact part is almost secondary and the topic stuff is really what is up. Topics are used below and it isn't even clear what they are or what we're talking about. Naming is important but lets tell the reader what is being named first.}

>>>>

% why are topics useful
Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifacts themselves.
Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}. 
Without the structured information which dashboards typically rely on, topics become a valuable and automatically extractable abstraction of software artifacts.
Topics abstract software artifacts by their subject matter, rather than their concrete representation. 
Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
The topics of development are relevant because they identify the distinct issues facing stakeholders at a particular time.

The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
% get neil's data mining point in there
To date, the results of topic analysis and concept location are project specific, and not generalizable.
% get the interpretation aspect
Much of the current research stops short of automatically interpreting the purpose of the topic, e.g.~\cite{Baldi2008,Kuhn2007,marcus04wcre}. 
That is, extracted topics are usually manually named and labelled by those who read the topic words.

% topics are numerous?
Deriving and labelling topics is a challenging machine learning problem. 
Topic analysis produces % raw topics that are represented as
 unlabelled context-free word-lists or word distributions.
Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
This manual approach is impractical when handling hundreds of different topics. To scale to large projects (thousands of lines of code and multiple releases), we need automatic assistance to determine the topic label.
We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.

@@

% harp on data mining more
In general, the mining of software artifacts tends to be very project specific, yet NFRs are not. 
% I was told if you're going to use a paren, you don't need to say it.
There is a series of standards on NFRs, such as \cite{iso9126}, specifically intended to apply to projects of varying types.
These standards indicate our approach is possible.%, if perhaps not universally accepted).
% NFRs are prevalent in almost all software systems.
% In this paper we will show that NFR related topics occur quite frequently.
%XXXXXX this sentence is confusing
%This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. 
In particular, we use software quality models to generalize topics across projects. 
This allows us to mine software systems with the intention of comparing their NFR-related development topics.
% in order to achieve generalizable, cross project data mining results.

@@

To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
We selected systems from the same application domain, so the functional requirements would be broadly similar. 
We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. 
The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
MySQL contains 320 thousand lines of C and C++ source code.  
Choosing an older version of these projects allowed us to focus on projects which have moved into the maintenance phase of the software life-cycle.

@@

For the commit message data-sets of each project, we created an XML file which separated commits into 30 day periods. 
This size of period, 30 days, is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze~\cite{Hindle09ICSM}. 
For each 30 day period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis algorithm~\cite{Blei2003}, and recorded the topics the algorithm returned.
%For each project and for each period of messages a project, we input these messages 
%we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.

@@

We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics~\cite{Hindle09ICSM}. 
% Furthermore, more than $20$ topics quickly became infeasible for inspection.

@@

\textbf{Do label frequencies change over time?} -- In both projects the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
As discussed in the previous section, certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time horizon being shorter than an equivalent span for the MySQL product. 
%%XXXXXXXXXXXXXXXXX what

@@

% para 4 summarize and hammer contribution
% NE not a fan of this phrasing -> % In conclusion, 
Previous topic analysis research produced topics that needed to be manually labelled, but topics without labels are difficult to interpret.
To improve this, we leveraged software engineering standards to produce a method of domain specific topic labelling.
Since the word-list technique is not project specific, it can be used to compare multiple projects.
We demonstrated multiple supervised and unsupervised methods of topic labelling, including multi-label learners.
We provided a case study and visualization of NFR related topics in MaxDB 7.500 and MySQL 3.23. 
During the case study we manually labelled hundreds of topics by hand and then used these annotations to validate the positive performance of topic labelling techniques.
The implication of our paper is that topics with labels are more useful for analyzing software development activities than are unlabelled topics. To furnish stakeholders with labelled topics, our technique provides many methods of automatic topic labelling.

@@

% Developers do more than merely observe these abstractions, however. They must also use them to assess progress and work with colleagues. To have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our development artifacts. Such labels can be used to compare progress or understand how well a project is meeting requirements. These labels then become the topics for development conversations (such as the stream of commentary on a bug report). One way to identify topics is to use the artifacts themselves to derive them. 
% \todo[inline]{it should be labels for topics not artifacts, I think the artifact part is almost secondary and the topic stuff is really what is up. Topics are used below and it isn't even clear what they are or what we're talking aout. Naming is important but lets tell the reader what is being named first.}
